---
title: 'Analysis '
author: "Anonymous"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  bookdown::pdf_book: default
geometry: margin=2cm
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 5,
  fig.align = 'center'
)
```

```{r load-libraries, include=FALSE}
rm(list = ls())
library(ggplot2)
library(cowplot)
library(dplyr)
library(PupillometryR)
library(ggpubr)
library(knitr)
library(ggforce)
library(tidyr)
library(rstatix)
```

```{r theme-setup, include=FALSE}
# Define conditions and visual parameters
CONDITIONS <- c('EA', 'TPEC', 'BO')
SHAPE <- c(21, 21, 21)
cb_palette <- c('#648FFF','#DC267F', '#FFB000')
TSIZE <- 17

# Theme for plots
p_theme <- theme(
  plot.title = element_text(face = "bold", hjust=0.5),
  panel.border = element_blank(),
  panel.grid.minor = element_blank(),
  axis.title = element_text(size=17),
  axis.text = element_text(size=11),
  axis.text.y = element_text(angle = 90, hjust = 0.5),
  legend.position = "none",
  panel.background = element_rect(fill = "#f1f2f5",
                                  colour = "white",
                                  size = 0.5, linetype = "solid")
)
```

```{r load-data, include=FALSE}
data_dir = '/Users/ethanhodess/Documents/Documents - Ethan’s MacBook Pro/Cedars/2025/TPOT_ensemble/collect/'
full_data <- read.csv(paste(data_dir, '500_imbalanced_balacc.csv', sep = "", collapse = NULL), header = TRUE, stringsAsFactors = FALSE)
```

TPOT best individual vs ensemble run with 500 pipeline evaluations. 

Top 5 imbalanced datasets from PMLB.

Ensemble initialized [] with greedy forward search across HDBSCAN cluster representatives and noise points.


# Task allbp

## Testing Balanced Accuracy
```{r task-allbp-test-plot, out.width="100%"}
plot_df <- full_data %>%
  filter(task.id == "allbp") %>%
  select(individual, ensemble) %>%
  pivot_longer(
    cols = c(individual, ensemble),
    names_to = "type",
    values_to = "accuracy"
  ) 

model_accuracy <- ggplot(
  plot_df,
  aes(x = type, y = accuracy, fill = type)
) +
  geom_flat_violin(
    position = position_nudge(x = 0.1),
    scale = "width",
    alpha = 0.2,
    width = 1.5
  ) +
  geom_boxplot(
    color = "black",
    width = 0.08,
    outlier.shape = NA,
    alpha = 0.0,
    size = 0.8,
    position = position_nudge(x = 0.15)
  ) +
  geom_point(
    position = position_jitter(width = 0.015, height = 0.0001),
    size = 1.0,
    alpha = 1.0
  ) +
  p_theme +
  labs(
    title = "Individual vs Ensemble (Task allbp)",
    x = "",
    y = "Accuracy"
  ) +
  theme(
    legend.position = "none",
    plot.title = element_text(hjust = 0.5)
  ) 

model_accuracy
```


# Task allhyper

## Testing Balanced Accuracy
```{r task-allhyper-test-plot, out.width="100%"}
plot_df <- full_data %>%
  filter(task.id == "allhyper") %>%
  select(individual, ensemble) %>%
  pivot_longer(
    cols = c(individual, ensemble),
    names_to = "type",
    values_to = "accuracy"
  ) 

model_accuracy <- ggplot(
  plot_df,
  aes(x = type, y = accuracy, fill = type)
) +
  geom_flat_violin(
    position = position_nudge(x = 0.1),
    scale = "width",
    alpha = 0.2,
    width = 1.5
  ) +
  geom_boxplot(
    color = "black",
    width = 0.08,
    outlier.shape = NA,
    alpha = 0.0,
    size = 0.8,
    position = position_nudge(x = 0.15)
  ) +
  geom_point(
    position = position_jitter(width = 0.015, height = 0.0001),
    size = 1.0,
    alpha = 1.0
  ) +
  p_theme +
  labs(
    title = "Individual vs Ensemble (Task allhyper)",
    x = "",
    y = "Accuracy"
  ) +
  theme(
    legend.position = "none",
    plot.title = element_text(hjust = 0.5)
  ) 

model_accuracy
```

# Task allrep

## Testing Balanced Accuracy
```{r task-allrep-test-plot, out.width="100%"}
plot_df <- full_data %>%
  filter(task.id == "allrep") %>%
  select(individual, ensemble) %>%
  pivot_longer(
    cols = c(individual, ensemble),
    names_to = "type",
    values_to = "accuracy"
  ) 

model_accuracy <- ggplot(
  plot_df,
  aes(x = type, y = accuracy, fill = type)
) +
  geom_flat_violin(
    position = position_nudge(x = 0.1),
    scale = "width",
    alpha = 0.2,
    width = 1.5
  ) +
  geom_boxplot(
    color = "black",
    width = 0.08,
    outlier.shape = NA,
    alpha = 0.0,
    size = 0.8,
    position = position_nudge(x = 0.15)
  ) +
  geom_point(
    position = position_jitter(width = 0.015, height = 0.0001),
    size = 1.0,
    alpha = 1.0
  ) +
  p_theme +
  labs(
    title = "Individual vs Ensemble (Task allrep)",
    x = "",
    y = "Accuracy"
  ) +
  theme(
    legend.position = "none",
    plot.title = element_text(hjust = 0.5)
  ) 

model_accuracy
```


# Task dis

## Testing Balanced Accuracy
```{r task-dis-test-plot, out.width="100%"}
plot_df <- full_data %>%
  filter(task.id == "dis") %>%
  select(individual, ensemble) %>%
  pivot_longer(
    cols = c(individual, ensemble),
    names_to = "type",
    values_to = "accuracy"
  ) 

model_accuracy <- ggplot(
  plot_df,
  aes(x = type, y = accuracy, fill = type)
) +
  geom_flat_violin(
    position = position_nudge(x = 0.1),
    scale = "width",
    alpha = 0.2,
    width = 1.5
  ) +
  geom_boxplot(
    color = "black",
    width = 0.08,
    outlier.shape = NA,
    alpha = 0.0,
    size = 0.8,
    position = position_nudge(x = 0.15)
  ) +
  geom_point(
    position = position_jitter(width = 0.015, height = 0.0001),
    size = 1.0,
    alpha = 1.0
  ) +
  p_theme +
  labs(
    title = "Individual vs Ensemble (Task dis)",
    x = "",
    y = "Accuracy"
  ) +
  theme(
    legend.position = "none",
    plot.title = element_text(hjust = 0.5)
  ) 

model_accuracy
```

# Task hypothyroid

## Testing Balanced Accuracy
```{r task-hypothyroid-test-plot, out.width="100%"}
plot_df <- full_data %>%
  filter(task.id == "hypothyroid") %>%
  select(individual, ensemble) %>%
  pivot_longer(
    cols = c(individual, ensemble),
    names_to = "type",
    values_to = "accuracy"
  ) 

model_accuracy <- ggplot(
  plot_df,
  aes(x = type, y = accuracy, fill = type)
) +
  geom_flat_violin(
    position = position_nudge(x = 0.1),
    scale = "width",
    alpha = 0.2,
    width = 1.5
  ) +
  geom_boxplot(
    color = "black",
    width = 0.08,
    outlier.shape = NA,
    alpha = 0.0,
    size = 0.8,
    position = position_nudge(x = 0.15)
  ) +
  geom_point(
    position = position_jitter(width = 0.015, height = 0.0001),
    size = 1.0,
    alpha = 1.0
  ) +
  p_theme +
  labs(
    title = "Individual vs Ensemble (Task hypothyroid)",
    x = "",
    y = "Accuracy"
  ) +
  theme(
    legend.position = "none",
    plot.title = element_text(hjust = 0.5)
  ) 

model_accuracy
```

\newpage
# Stats Testing
```{r stats, out.width="100%"}
# Wilcoxon rank-sum test
results <- full_data %>%
  pivot_longer(
    cols = c(individual, ensemble),
    names_to = "type",
    values_to = "accuracy"
  ) %>%
  group_by(task.id) %>%
  wilcox_test(accuracy ~ type, paired = TRUE) %>%
  adjust_pvalue(method = "holm") %>%
  add_significance()

print(results)
```



